Inferring disease architecture and predictive ability\\with LDpred2-auto

### 2023-10-15 paper/paper-infer.tex
minor changes

Across 169 UKBB phenotypes with more than 25 chains kept, we compute the median heritability per block, and compare it to the number of variants in these blocks; the median heritability explained by a block of variants is largely proportional to the number of variants in this block (Figure S38).
The outlier block explaining a much larger heritability contains the HLA region.
Across the same phenotypes, we then compute per-variant median probabilities of being causal, and report them in a Manhattan-like plot in Figure S39.
Some variants in multiple small regions across the genome have a larger probability of being causal across many phenotypes; interestingly, these are mapped to genes that are known to be associated with many different traits (up to more than 300) in the GWAS Catalog \cite[]{buniello2019nhgri}.
To verify that this is not driven by population structure, we compute pcadapt chi-square statistics that quantify whether a variant is associated with population structure \cite[]{prive2020performing}; the log-statistics have a small negative correlation of -5.5\% with the probabilities of being causal.
To verify that this does not correspond to regions of low LD, we compute LD scores; the median probabilities of being causal have a small correlation of 11.6\% with the LD scores and of -6.8\% with the log of LD scores.

>>>>

Across 169 UKBB phenotypes with more than 25 chains kept, we compute the median heritability per block, and compare it to the number of variants in these blocks; the median heritability explained by a block of variants is largely proportional to the number of variants in this block (Figure S38).
The outlier block explaining a much larger heritability contains the HLA region.
Across the same phenotypes, we then compute per-variant median probabilities of being causal, and report them in a Manhattan-like plot in Figure S39.
Some variants in multiple small regions across the genome have a larger probability of being causal across many phenotypes; interestingly, these are mapped to genes that are known to be associated with many different traits (up to more than 300) in the GWAS Catalog \cite[]{buniello2019nhgri}.
To verify that this is not driven by population structure, we compute pcadapt chi-square statistics that quantify whether a variant is associated with population structure \cite[]{prive2020performing}; the log-statistics have only a small correlation of 5.9\% with the probabilities of being causal.
To verify that this does not correspond to regions of low LD, we compute LD scores; the median probabilities of being causal have a correlation of 32.0\% with the LD scores and of 26.4\% with the log of LD scores.
Therefore the posterior probabilities of being causal obtained from LDpred2-auto tend to slightly increase with LD scores.
